### 2025-05-23 manuscript/supplemental/supplemental.qmd

---
title: Generating Correlated Data for Omics Simulation
author:
    - name: "Jianing Yang"
      orcid: 0000-0002-2048-9398
      affiliations:
        - name: "Chronobiology and Sleep Institute, University of Pennsylvania"
    - name: "Gregory R. Grant"
      orcid: 0000-0002-0139-7658
      affiliations:
        - name: "Department of Genetics, University of Pennsylvania"
    - name: "Thomas G. Brooks"
      orcid: 0000-0002-6980-0079
      affiliations:
        - name: "Institute for Translational Medicine and Therapeutics, University of Pennsylvania"
editor: 
  markdown: 
    wrap: sentence
execute:
  echo: false
  cache: true
  warning: false
format:
  html:
    toc: true
    toc-expand: true
bibliography: ../references.bib
---

## Supplemental Material

### Comparisons to real data

```{r}
#| label: fig-compare-to-real
#| fig-cap: "Comparison to real data run on a fly whole body data set from GSE81142. (a-b) Comparison of gene (a) mean expression and (b) variance, log-scaled in real and PCA simulated data. The line of equality is marked in black. Points are colored according to the density of points in their region. Wishart and corpcor methods give similar results (not shown). (c) Quantile-quantile plot comparing correlation values of gene pairs from real data and simulated data (both with and without dependence). Genes with at least 30 reads were used. Values on the diagonal line indicate a match between the simulated and real data sets. (d) Projections onto the top two principal components of the real data set for both real and simulated data. All 8 simulations (96 samples for each simulation) shown. (e) Principal component analysis was performed on all data sets and the variance captured by the top components is shown. Unlike (d), these components were fit from each data set considered separately instead of reusing the weights from the real data."
#| fig-width: 10
#| fig-height: 10
readRDS("../../processed/compare_to_real_plot.GSE81142.RDS")
```

```{r}
#| label: fig-compare-to-real2
#| fig-cap: "Comparison to real data run on a mouse cortex timeseries body data set from GSE151565. Only timepoint ZT0 was used for this comparison. (a-b) Comparison of gene (a) mean expression and (b) variance, log-scaled in real and PCA simulated data. The line of equality is marked in black. Points are colored according to the density of points in their region. Wishart and corpcor methods give similar results (not shown). (c) Quantile-quantile plot comparing correlation values of gene pairs from real data and simulated data (both with and without dependence). Genes with at least 30 reads were used. Values on the diagonal line indicate a match between the simulated and real data sets. (d) Projections onto the top two principal components of the real data set for both real and simulated data. All 8 simulations (96 samples for each simulation) shown. (e) Principal component analysis was performed on all data sets and the variance captured by the top components is shown. Unlike (d), these components were fit from each data set considered separately instead of reusing the weights from the real data."
#| fig-width: 10
#| fig-height: 10
readRDS("../../processed/compare_to_real_plot.GSE151565.RDS")
```

## Supplemental Methods

Here, we provide an introduction to the Gaussian copula approach.

### Multivariate normal distribution

We first discuss the simplest case, where our dataset is multivariate normally distributed.
The distribution $N(\mu, \Sigma)$ is the multivariate normal distribution with mean vector $\mu$ and covariance matrix $\Sigma$.
Here $\mu$ is a $p \times 1$ column vector and $\Sigma$ is a $p \times p$ matrix.
In order for this to work, $\Sigma$ must be symmetric and positive semi-definite, meaning that all of its eigenvalues are non-negative. 
This matrix is conceptually simple, since $\Sigma_{ij}$ gives the covariance of $x_i$ and $x_j$ when $x \sim N(\mu, \Sigma)$.
More specifically, this is the population covariance matrix of $N(\mu, \Sigma)$, which does not generally equal the sample covariance matrix.
Indeed, if $X$ is $n$ samples from $N(\mu, \Sigma)$, then $\widehat \Sigma := (X - \bar X) (X - \bar X)^T / (n - 1)$ is the sample covariance matrix where $\bar X$ is the average of the $n$ samples.
For large $n$, $\widehat \Sigma$ will closely approximate $\Sigma$, but we care primarily about the situation where $n$ is small.
In particular, $\widehat \Sigma$ is at most a rank $n - 1$ matrix while $\Sigma$ could be up to rank $p$.
This means that $N(\mu, \widehat \Sigma)$ and $N(\mu, \Sigma)$ are quite different distributions: every sample from $N(\mu, \hat \Sigma)$ is contained in an $n-1$ dimensional plane.
If $n$ is small, then this is very unlike real data, which typically is close to $p$ dimensional.

The difficulty then is that we only know $\widehat \Sigma_{ref}$ from our reference dataset, but we need to choose a $\Sigma_{sim}$ with which we can simulate data and the obvious choice of $\widehat \Sigma_{ref}$ is inadequate.
The most common choice is to assume $\Sigma_{sim}$ is a diagonal matrix.
This is the situation we want to avoid where the generated data is independent: $x_i$ and $x_j$ have zero covariance unless $i = j$.
However, this has some nice properties, such as being simple and fast to simulate (just generate univariate normal data for each variable).

We describe three alternative approaches in Methods of how to choose a $\Sigma_{sim}$ that will produce data similar to the input data.
All three of these rely on a specific form of $\Sigma$, namely that
$$ \Sigma_{sim} = D^2 + UW^2U^T $$
where $D$ is a $p \times p$ diagonal matrix, $U$ is $p \times k$ for some $k \ll p$ and $W$ is a diagonal $k \times k$ matrix.
This is a combination of an independent part (the diagonal matrix) and a low-rank part ($U W^2 U^T$ is rank at most $k$).

Generating data for $\Sigma_{sim}$ of this form is highly efficient.
We use two basic facts about the multivariate distribution.
First, if $A$ is a matrix and $u ~ N(0, \Sigma)$ then $Au ~ N(0, A \Sigma A^T)$.
Second, if $u ~ N(0, \Sigma_1)$ and $v ~ N(0, \Sigma_2)$, then $u + v ~ N(0, \Sigma_1 + \Sigma_2)$.
Therefore, given matrices $D, U, W$ and $u ~ N(0, I_k)$ and $v ~ N(0, I_p)$ (where $I_k$ and $I_p$ are the $k \times k$ and $p \times p$ identity matrices), then
$$ D v + U W u ~ N(0, D^2 + UW^2 U^T) = N(0, \Sigma) $$
as desired.

In contrast, when given an arbitray $\Sigma$, to generate a random value in $N(0, \Sigma)$, one must compute a matrix $V$ such that $VV^T = W$.
Then, using $v ~ N(0,I)$ , we have $Vv ~ N(0, \Sigma)$ as desired.
However, computing $V$ is done using a Cholesky decomposition or eigenvector decomposition, both of which are computationally expensive when $\Sigma$ is large.

Lastly, we emphasize that multivariate normal distributions do not capture all, or even most, types of possible dependence.
Indeed, we see this even in the 2-dimensional case where it is well known that correlation describes only a linear relationship between two variables while in reality they may have much more complex relations.
In higher dimensions, the problem is only worse.
So any method based off multivariate normal distributions are making large assumptions about distribution.
However, it is necessary to make some assumption like this.
In the next section, though, we see that "normal" part is actually not a large obstacle.

### Gaussian copula

Building on the multivariate normal distribution, a popular approach to describe dependence in a high-dimensional settings is called the Gaussian copula approach.
The idea of this approach is that by applying a normalizing transform and later reversing the transformation, data that does not fit a normal distribution can still have its dependence structure described using a multivariate normal distribution.
This allows the marginal (i.e., univariate) distributions of each genes to be specified separately from the dependence between genes.
This operates first by transforming each gene by fitting a distribution (such as a normal distribution, Poisson, negative binomial, or other form), and then applying the fit cumulative distribution function (CDF) to the observed values.
Finally, those are fed to a standard normal distribution's inverse CDF to obtain values that are approximately normally distributed.
These values are then used to compute a covariance matrix $\Sigma$ and the data is assumed to follow a multivariate normal distribution in $p$ dimensions with that covariance matrix.

Here, we describe the approach using the form of covariance matrix $\Sigma = D^2 + UW^2U^T$ as above.
Once data is obtained $Z \sim N(0, \Sigma)$, then one can undo the normalizing transformation to obtain data with the same marginal distributions as the fit marginal distributions but with dependence determined by $\Sigma$.
We describe this in detail:

1.  Fit marginal distributions to each feature in $X$ to determine CDFs $F_{i}$ for each feature.
2.  Apply normalizing transform to $X$ by setting $Z_{ij} = \Phi^{-1}(F_{i}(X_{ij}))$ where $\Phi$ is the CDF of the standard normal distribution.
3.  Compute $D$, $U$, $W$ matrices from $X$ by one of three methods (see Methods).
4.  Generate $k$ i.i.d. standard normally distributed values $u$ and $p$ i.i.d standard normally distributed values $v$.
5.  Set $z' = UWu + D v$.
6.  Output the vector $x'$ where $x'_i = F_i^{-1}(\Phi(z'))$.

The generated data vector $z'$ has covariance matrix $\Sigma_{sim} = D^2 + U W^2 U^T$.
Moreover, we require that $\Sigma$ satisfies that $\left(\Sigma_{sim}\right)_{ii}$ is approximately 1 for each $i$.
That guarantees that the output $x'$ has each entry with the same marginal distributions $F_i$ as was originally fit and inherits gene-gene dependence from $Z'$.
This method is computationally efficient, taking hardly any more time or memory than simulations without dependence.

### Running SPsimSeq

The R package SPsimSeq [@Assefa2020] provides a dedicated RNA-seq and single-cell RNA-seq simulator using a Gaussian copula approach to simulate gene dependence.
We ran SPsimSeq using the following options:

``` R
    n.sim = 1,
    s.data = input,
    n.genes = input |> nrow(),
    batch.config = 1,
    group.config = 1,
    pDE = 0,
    tot.samples = 100,
    model.zero.prob = FALSE,
    genewiseCor = TRUE,
    log.CPM.transform = TRUE,
    result.format = "list",
    return.details = TRUE,
    verbose = TRUE,
```

This produces bulk RNA-seq data with no batching or group effects or differentially expressed (DE) genes.
Raw count values provided as the input matrix.
For dataset GSE151565, only the ZT0 timepoints were used.
This is the same set of samples that were used in our simulation to determine the correlation structure (copula).

SPsimSeq was not run in the two example applications, DESeq2 and CYCLOPS, due to substantial differences in how it generates DE across samples compared to our method.
DE in SPsimSeq is done by mimicking a random set of detected DE genes in the reference data set.
SPsimSeq does not report either the group mean values or the log fold changes used for these genes, so we are not able to modify our own simulation to have the same set of DE genes at the same effect sizes.

### Running vine copula

We used the rvinecopulib package to fit and generate random data from a vine copula distribution.
This package is a R wrapper of the high-performance vinecopulib C++ package.
This package expects data normalized to the range of \[0, 1\].
Therefore, we chose not to use the real RNA-seq count data and instead generated correlated random number as input which follow a multivariate normal distribution with covariance matrix constructed with a randomly generated singular vector matrix and singular values 1, 2, ..., 100.
These values were then normalized to the interval \[0,1\] by the `pnorm` function.
A vine copula was fit using just the "guassian" family.
While rvinecopula supports multi-threaded operation, we ran using a single thread as with the other methods.

### Running mvrnorm

To assess a simple multivariate normal distribution simulation option, we ran the `mvrnorm` function from the MASS R package.
We set means to be the mean expression of each gene and used the sample covariance matrix as the `Sigma` parameter.